---
title: "Practical Work"
author: "Raul Gonzalez Duarte"
date: "30/1/2022"
output: html_document
---

## Song analysis

### Setup and Load data

```{r setup}
# Install required packages if needed
source("./requirements.R")

# Libraries
library(tm)
library(textclean)
library(dplyr)
library(ggplot2)
library(ggwordcloud)
library(udpipe)
library(kableExtra)
library(sentimentr)
library(SentimentAnalysis)

# Load data
df <- read.csv('data/DuaLipa.csv')

# Remove id and artist column
df$X <- NULL
df$Artist <- NULL

# Last released album
df2 <- filter(df,Album == "Future Nostalgia")
df2[1,1]= "Don't Start Now" # Title fix
```


### Album

```{r album data cleaning}
# Read data
CORPUS <- VCorpus(VectorSource(df2$Lyric))

# Data cleaning 
CORPUS <- tm_map(CORPUS, content_transformer(tolower))
CORPUS <- tm_map(CORPUS, content_transformer( function(x) gsub(" ", " ", x))) # song's fix
CORPUS <- tm_map(CORPUS, content_transformer(function(x){replace_contraction(x)}))
CORPUS <- tm_map(CORPUS, content_transformer(removePunctuation), ucp = F)
CORPUS <- tm_map(CORPUS, content_transformer(removeNumbers))
CORPUS <- tm_map(CORPUS, removeWords, stopwords("english"))
CORPUS <- tm_map(CORPUS, stripWhitespace)

```

Now we create the term document matrix and with that, a dataframe of frequent terms is created and used to plot a word cloud
```{r wordcloud}
tdmTR <- TermDocumentMatrix(CORPUS)

freq <- findFreqTerms(tdmTR, lowfreq = 1, highfreq = Inf)
freq <- as.matrix(tdmTR[freq,])
freq <- as.data.frame(rowSums(freq))
colnames(freq) <- "num"
freq$word <- rownames(freq)

options(repr.plot.width = 18, repr.plot.height = 8) 
set.seed(5555)
freq$angle <- 45 * sample(-2:2, nrow(freq), replace = TRUE, prob = c(1, 0, 4, 0, 1))

ggplot(freq, aes(label = word,size = num, color = num, angle = angle)) +
  geom_text_wordcloud(shape = "circle", rm_outside = TRUE, area_corr = F, rstep = .01,max_grid_size = 256,grid_size = 7,grid_margin = .4) +
  scale_size_area(max_size = 12.5) + theme_minimal() + scale_color_gradient(low = "darkgrey", high = "#53d1b1")
```

Number of words in the album
```{r nWords}
paste("Total words:", sum(freq$num), "; Unique words:",nrow(freq))
```

With the tdm, sort the words by frequence and create a horizontal bar plot with the top 10 words of the album
```{r topWords}
matrix <- as.matrix(tdmTR) 
words <- sort(rowSums(matrix),decreasing=TRUE) 
df3 <- data.frame(freq=words[1:10],word = names(words[1:10]))
barplot(df3[order(df3[,1],decreasing=FALSE),][,1],names.arg=df3[order(df3[,1],decreasing=FALSE),][,2]
        ,col="green",xlab="Frequency",ylab="Word",main="Top 10 Album words",border="black",horiz = TRUE)

```

Now with the udpipe, create a dataframe with the types an number of words of each type in the album
```{r typeWords}
dl <- udpipe_download_model(language = "english")
udmodel_eng <- udpipe_load_model(file = "english-ewt-ud-2.5-191206.udpipe")
x <- udpipe_annotate(udmodel_eng, x = df2$Lyric)
x <- as.data.frame(x)
table(x$upos)
```

#### Plot it with library kableExtra
```{r typeWordsPlot}
#The first 4 cols UNSHOWN are doc_id, paragraph_id, sentence_id and sentence
kable_styling(kable(x[1:20, c(5:9)]),font_size = 7)
kable_styling(kable(x[1:20, c(10:14)]),font_size = 7) #Remaining cols

```


### Song classification

Now we use two libraries to classify the songs. Sentiments less than zero is negative, 0 is neutral, and greater than zero positive polarity
```{r sentiments}
s1 <- sentiment(df2$Lyric)

sentiment <- analyzeSentiment(df2$Lyric)

song <- data.frame(Title=df2$Title,sentimentr=s1$sentiment,SentimentAnalysis=sentiment$SentimentQDAP,
                 SentimentAnalysisDir=convertToDirection(sentiment$SentimentQDAP))
song
```

### Individual song analysis
For each song, we calculate the word cloud, the total and unique number of words and the top 10 frequent words.
```{r}
# WordCloud, number of words and top 10 by song
j<-1
for (i in df2$Lyric) {
  print(paste("Song:",df2$Title[j]))
  
  # Read data
  CORPUSSong <- VCorpus(VectorSource(i))
  
  # Cleaning 
  CORPUSSong <- tm_map(CORPUSSong, content_transformer(tolower))
  CORPUSSong <- tm_map(CORPUSSong, content_transformer( function(x) gsub(" ", " ", x))) # song's fix
  CORPUSSong <- tm_map(CORPUSSong, content_transformer(function(x){replace_contraction(x)}))
  CORPUSSong <- tm_map(CORPUSSong, content_transformer(removePunctuation), ucp = F)
  CORPUSSong <- tm_map(CORPUSSong, content_transformer(removeNumbers))
  CORPUSSong <- tm_map(CORPUSSong, removeWords, stopwords("english"))
  CORPUSSong <- tm_map(CORPUSSong, stripWhitespace)
  
  
  # WordCloud by song
  tdmTRSong<- TermDocumentMatrix(CORPUSSong)

  freqSong <- findFreqTerms(tdmTRSong, lowfreq = 1, highfreq = Inf)
  freqSong <- as.matrix(tdmTRSong[freqSong,])
  freqSong <- as.data.frame(rowSums(freqSong))
  colnames(freqSong) <- "num"
  freqSong$word <- rownames(freqSong)
  

  options(repr.plot.width = 18, repr.plot.height = 8) 
  set.seed(5555)
  freqSong$angle <- 45 * sample(-2:2, nrow(freqSong), replace = TRUE, prob = c(1, 0, 4, 0, 1))

  print(ggplot(freqSong, aes(label = word,size = num, color = num, angle = angle)) +
    geom_text_wordcloud(shape = "circle", rm_outside = TRUE, area_corr = F,rstep = .01,max_grid_size = 256,grid_size = 7,grid_margin = .4) +
    scale_size_area(max_size = 12.5) + theme_minimal() + scale_color_gradient(low = "darkgrey", high = "#53d1b1"))
  
  # Number of words in the song
   print(paste("Total words:", sum(freqSong$num), "; Unique words:",nrow(freqSong)))
   
  # top 10 song words
  matrixSong <- as.matrix(tdmTRSong) 
  wordsSong <- sort(rowSums(matrixSong),decreasing=TRUE) 
  dfSong <- data.frame(freq=wordsSong[1:10],word = names(wordsSong[1:10]))
  barplot(dfSong[order(dfSong[,1],decreasing=FALSE),][,1],names.arg=dfSong[order(dfSong[,1],decreasing=FALSE),][,2]
          ,col="blue",xlab="Frequency",ylab="Word",main="Top 10 Song words",border="black",horiz = FALSE)
  j <- j+1
}
```


